Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-29 ba_resubmit/response_to_reviewers3.tex

%Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.

\section{AE Comments}

\pointRaised{AE}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}

\reply{We have made several revisions to make the paper easier to read, as described in our responses below. 
}

@@

\pointRaised{AE}{Line 80 on p. 3 says: “For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say “record $i \in X_1$” rather than the more compact $x_{1i}$.” Taking this at face value implies that, in any sentence, one could swap “record $i \in X_1$” for “$x_{1i}$,” which clearly is not true.  So, as it stands, the sentence does little to improve ease of readability.  In fact, if the point is that “$x_{1i}$” will never be used again, why is the notation introduced in the first place?}

%%% RCS: Thank the Editor, explain the change. I have removed repetition and shortened this bit.

\reply{
To lighten the notation and allow for greater precision, we now refer to the two data files as $A$ and $B$. The updated material now reads:

``Consider two data files $A$ and $B$ comprising $n_A$ and $n_B$ records, respectively, and including $F$ linkage variables measured in both files. For $i=1, \dots, n_A$, let record $i$ be given by $A_i = (A_{i1}, \dots, A_{iF})$, so that $A = (A_i : i = 1, \dots, n_A)$.  Similarly, for $j=1, \dots, n_B$, let record $j$ be given by $B_j = (B_{j1}, \dots, B_{jF})$, so that $B = (B_j : j = 1, \dots, n_B)$.  Without loss of generality, denote files such that $n_A \geq n_B$.". See Section 2. 
}

@@

\setcounter{responsectr}{0}
\section{Reviewer Comments}

	\pointRaised{R}{%
	For the missing data treatment now included just before Section 2.1, is the assumption truly missing at random, or missing completely at random? Or, does this distinction not matter because of the independence assumed across elements of the comparison vector?}

	\reply{Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here. For simplicity, we have chosen to use MCAR in the article. We have included the following explanation of missingness in Section 2.

	``In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}). To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data."
	}

@@

\reply{%
	The square brackets have been corrected.

	We have revised the derivation of the full conditional distribution of $Z_j$, separately handling the case when $B_j$ has a match and when it does not. Refer to Appendix A for this extended derivation.
	}

	\pointRaised{R}{%
Appendix 8.2: Thank you for including the details on integrating out pi from the full conditionals. They surprised me. I had assumed that the authors had integrated out pi in the prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite matchings”. (1) Does the alternate approach presented in this paper provide a different algorithm than directly integrating the prior distribution? (2) Is the presented alternative approach justified?}

\reply{%
		We realize that the integration of $\pi$ out of the full conditionals is not necessary for our sampler. Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler.

		We have redone all simulations and case studies using a Gibbs sampler that samples $\pi$ from its full conditional, and then samples $\bm{Z}$ from its full conditionals. We obtained equivalent results.
	}

%\reply{No, in \text{fabl}, we do not integrate out $\pi$ as done in \cite{sadinle_bayesian_2017} because this leads to a sequential sampler. 
%
%(1) Yes, our alternative approach is different than directly integrating over the prior distribution $\pi.$ 
%%We have attempted to make this more clear in our revised version. 
%In short, our prior distribution over $\bm{Z}$ leads to a Gibbs sampler that leads to parallel updates, which contrasts that of \cite{sadinle_bayesian_2017}. Of course, we could integrate out $\pi$ in our approach, however, this leads to sequential updates, which would be slow in practice. See Appendix A, page 27. 
%
%(2) Yes, our alternative approach is justified. We provide a justification in Appendix A, page 27 regarding why the updates are independent and not sequential. Thank you for the excellent questions, which have greatly improved our paper. 
%
%}

@@

\pointRaised{R}{%
	I do not understand the statement just below Equation (10): “When j has no match in $X_1$, we write $(n_1 +j, j) \in h_{P+1}$” My understanding of these patterns is that they are based on observed comparison vectors without consideration of Z (matches). In the second paragraph of Section 4.3, the H notation includes the matches (Z), in notation and definition that seems to conflict with the statement just below Equation (10). This also comes into play in Equation (16)}

	%% RCS: Tried to answer this from the original paper and the updated paper. Tried to make it shorter and close it out very quickly.

	\reply{Yes, the comparison vectors are created (and therefore, patterns are assigned) without regard to $\bm{Z}$. We have removed that sentence, and updated the equations in the Section 4.3 to avoid confusion}

@@

\pointRaised{R}{%
	Third paragraph of Section 4.1: The authors claim they are computing “sufficient statistics”. What exactly are these statistics sufficient for?
	}

	\reply{We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$. However, we cannot express the full likelihood in (14) through these statistics. Therefore, we have revised to text to say ``summary statistics" instead of  ``sufficient statistics."

	%Thank for pointing out our typo. We have revised the text to state ``summary statistics" instead of ``sufficient statistics."
	}

@@

\pointRaised{R}{%
	First paragraph of Section 4.2: The authors state: “Posterior calculations still attribute the appropriate weight to all records through the summary statistics…” What is meant by the term “weight”? Which records are appropriately weighted – those in X2?}

		%% Comment: Brian, you have not answered the reviewer's question regarding what is meant by weight and what records are appropriately weighted. \textcolor{red}{Please revise to directly answer the question in the revised text with equation numbers. To my understanding, the weights have been removed and section 4 has been rewritten }}
%	\reply{
%	\textcolor{teal}{In the original manuscript, we consider all records in the first file ($X_{1}$) that share an agreement pattern with the second data file ($X_{2}$). In this situation, all these records have the same Fellegi-Sunter likelihood ratio weight, $w_{ij}$. Given this result, we denoted the weight $w_p$ since these records map to the same agreement pattern.} 
%	
%	\textcolor{teal}{In the revised manuscript, we define $m_p$ and $u_p,$ which are the 
%	 probabilities that records $A_i$ and $B_j$ form agreement pattern $p$ given that
% they are a match and non-match, respectively. For each pattern $p,$ we define $w_p = m_p/u_p.$ We hope that this will be more clear than our previous presentation of the material. See Section 4.1, pages 8--9.}
% }
% 
	\reply{Our intended meaning was ``All record pairs contribute to the likelihood through the summary statistics."

	In our revised submission, we have reexpressed the likelihood from (6a) through the perspective of agreement patterns in (14). This form of the likelihood is then used to express contributions to the likelihood for the $\bm{m}$ and $\bm{u}$ parameters in (15) and for $\bm{Z}$ in (17), both of which use the summary statistics. We hope that this revision makes this point more clear.}

@@

\pointRaised{R}{%
	Second paragraph of Section 4.2: “and delete those comparison vectors”. Which are “those” vectors?}

	\reply{

%	In the revised paper, ``vector" is replaced with ``matrix" when referring to $\gamma.$ (See R.13 regarding making the paper consistent below).

	In the original manuscript, ``those comparison vectors" refer to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$

	The revised text states: ``Then, we conduct hashing, obtain the compressed comparison matrix, $\tilde{\gamma}^{ab}$, and remove the memory-intensive comparison matrix, $\gamma^{ab}$, before continuing with the next chunk of data. See Section 4.3}

%	 \textcolor{red}{Question for Brian: is lowercase $\gamma$ always a matrix? Let's make sure we're not swapping between vectors and matrices. In the previous version this was a vector and now its listed as a matrix. I would argue that a matrix should use capital notation, such as $\Gamma.$}

@@

%\reply{%
%	We have revised that sentence to read ''We then conduct hashing, obtain the compressed $\tilde{\Gamma}^{ab}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data."
%}

\pointRaised{R}{%
	Where does $R^{SEI, cd}$ come into play in the partitioned algorithm presented in Equations (13) and (14)? I recommend that the authors either refrain from suppressing the SEI notation or further explain how the SEI algorithm has changed the quantities in these equations.}

\reply{To make the SEI algorithm more clear, we have revised the ordering of Sections 4, so that we have 1) hashing, 2) posterior inference, 3) chunkwise computation of the comparison matrix, and 4) SEI. With this new ordering, and with the newly provided equations (15) through (19), we hope it clearer that that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$. In particular, (16a), (16b), and (18) depend only $\mathcal{N}$. The SEI algorithm only affects the step shown in (19). 
	}

@@

%\reply{%
%	We have reordered Section 4 so that all posterior inference is presented before SEI. This makes it clear that the posterior updates in (16a), (16b), and (18) depend only on $\mathcal{N}$. SEI only affects the step shown in (19).
%}

\pointRaised{R}{%
	Section 4.2: I appreciate the practical advice about choosing S for the SEI method. However, this choice seems arbitrary in the absence of further discussion/evidence. Given that the primary novelty of the manuscript is in methods to speed and otherwise improve computation, I am surprised that this aspect of computational innovation is presented with virtually no theoretical or empirical exploration. Presumably the SEI method has some sort of accuracy trade-off, as the authors warn that linkage results may be “distorted” if S is low. However, this trade-off is not quantified or even discussed in practical terms beyond the terse recommendation to choose S=10.
}

\reply{% 
	We have included an additional simulation in Section 5.3, which explores the trade-offs regarding different choices of $S$. 
}
\pointRaised{R}{%
	Page 5, 2 sentences before equation (4): I believe the sum should be of $I(Z_j \leq n_1)$, not $I(Z_j \leq n_1 +1)$.}

\reply{%
	This has been corrected.
}

@@

\pointRaised{R}{%
	Equation (6a): The indices do not match the subscripts in the indicator function in each summand, or their standard meaning in table 1.}

\reply{%
	The typo has been fixed.
}

\pointRaised{R}{%
	Generally, the authors seem to arbitrarily use upper and lower case z interchangeably in function definitions.}

\reply{%
	We use $\bm{Z}$ when discussing a random quantity. In all instances where we discuss $Z_j$ taking on a particular value, we now use $q$ to avoid confusion and to follow notation used by \cite{sadinle_bayesian_2017}.
}

@@

\pointRaised{R}{%
Equations (8) and (9): Should the weights have superscript (s) (as the Zs do)?}

\reply{%
	This has been fixed. See Equations (9) and (10).
}

\pointRaised{R}{%
	Gamma is in some places described as a set and in others as a matrix (particularly in sections 4.2 and 5.1).
}

\reply{%
	$\gamma$ is a comparison matrix comprised by comparison vectors. All references to $\gamma$ as a set have been removed. See Section 2, where we define both the comparison vector and comparison matrix.
}

### 2023-07-29 ba_resubmit/kundinger_resubmit2.tex
lots and lots of edits

\usepackage{lineno}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
\usepackage{graphicx}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{rem}{Remark}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}
\usepackage{subcaption}

@@

\begin{abstract} 
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, contrasting the sequential updates of \texttt{BRL}.}
\textcolor{teal}{We ensure our method is bipartite using decision theoretic post-processing techniques from the existing literature.}
\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead. Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing. 
}
%This allows
%us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. 
Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
\end{abstract}

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity. However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989} \textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}

@@

Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define a \emph{comparison vector} $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$. For notational convenience, we define the \emph{comparison matrix} $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.

@@

\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.  He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.  As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.  Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}

@@

\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors. To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR). Due the the conditional independence assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}) and perform all computation using the observed data. For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}

%Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}). With either the ignorability or MAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data. For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}. Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p. 12]{LittleRubin2002}). To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation.}

@@

\begin{rem}
In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$. This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$. In our preliminary studies, the this highly informative prior weighting on matching can result in overly optimistic match rates. Hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
\end{rem}

\begin{rem}
Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches. In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}). In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows. Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence. 
\end{rem}

@@

Next, the full conditional for $\pi$ is given by
\begin{align}
\label{eqn:pi}
	\pi^{(s+1)}|\bm{Z}^{(s)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
\end{align}
where $n_{AB}(\bm{Z}^{(s)}) = \sum_{j=1}^{n_B} I(Z_j^{(s)} \leq n_A)$.

@@

The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_A  n_B$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors. One approach to reduce the number of comparisons is blocking, which places similar records into partitions, or ``blocks'' \citep{christen2019data}. In deterministic blocking, the modeler chooses fields thought to be highly reliable and only compares records that agree on those fields. The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains. Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking \textcolor{teal}{often}
 undesirable %in some situations 
\citep{steorts_comparison_2014}.

@@

With \texttt{fabl}, we introduce two techniques to expand the scalability of probabilistic record linkage. First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors. Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.

@@

Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$. Let $P$ be the number of patterns realized in $\gamma$. 
%It is
\textcolor{teal}{The comparison vector} is
 bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1).$
% This comment below seems obvious and not needed.  
% where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values. 
\textcolor{teal}{The quantity $P^{*}$ clearly depends only on $F$ and $L_f.$ Observe it does not depend (or scale) with $n_A$ or $n_B.$}
%is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$.

To obtain a memory efficient representation, we map the agreement pattern of each record pair to a unique integer. \cite{enamorado2019using} accomplished this through a hashing function, which we modify to explicitly handle missing values:
\begin{align}
	\label{eqn:hashing}
	h^{*}(\gamma_{ij}) = \sum_{f = 1}^F I_{obs}(\gamma_{ij}^f)2^{\gamma_{ij}^f + I(f>1)\sum_{d=1}^{f-1}(L_d)}.
\end{align}

%\textcolor{red}{Do you mean what I have below or something else? This still is a bit hard to parse.}
We then map \textcolor{teal}{these integers in (\ref{eqn:hashing})} to sequential integers $\{1, \ldots, P\}$. \textcolor{teal}{Denote} each unique agreement pattern as $h_p$, and the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$. When record pair, \textcolor{teal}{$(i,j)$,} exhibits agreement pattern $p$, \textcolor{teal}{let} $\gamma_{ij} = h_p$.

In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$. This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.

\begin{example}
\label{ex:one-hot}
For example, with five fields with binary agreements and possible missingness (NA), the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$. Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, \text{NA}, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field. Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$. We map the unique values of $h^{*}(\gamma)$ to sequential integers. \textcolor{teal}{Suppose that 1066 maps to 42. This implies $\gamma_{5,7} = h_{42}.$}
%If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ 
This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.
\end{example}

@@

	\textcolor{red}{Suggestion: Chunk is award. Perhaps partition/block/batch/segment. I might like segment since partition/block have meanings in record linkage already. }

		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks for more manageable computations. Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.

		Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B$ and $B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.  For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records. For each $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.

>>>>

\usepackage{lineno}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
\usepackage{graphicx}
\newtheorem{lemma}{Lemma}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}
\usepackage{subcaption}

@@

\begin{abstract} 
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
\end{abstract}

%\textcolor{teal}{which allows for parallel updates to the Gibbs sampler, contrasting the sequential updates of \texttt{BRL}.}
%\textcolor{teal}{We ensure our method is bipartite using decision theoretic post-processing techniques from the existing literature.}
%\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead. Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing. 
%}

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity. However, in many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989}.
%\textcolor{teal}{or using decision theoretic procedures, such as those of \cite{sadinle_bayesian_2017} and \cite{steorts_bayesian_2016}, which both ensure  bipartite (or rather transitive) matches.}

@@

Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$. For notational convenience, we define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.

@@

\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.  He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.  As the prior distriburion for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.  Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}

@@

\textcolor{teal}{We now explain how we accomodate missing fields, which leads to missing comparison vectors. To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are missing completely at random (MCAR, per \cite{LittleRubin2002}). With the MCAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data.}

@@

\begin{remark}
In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$. This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$. In our preliminary studies, the this highly informative prior weighting on matching can result in overly optimistic match rates. Hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
\end{remark}

\begin{remark}
Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches. In the bipartite case, we conjecture that some analysts will find $\pi$ in (\ref{eqn:fast_beta_prior2}) to be more a interpretable parameter than $\lambda$ in (\ref{eqn:lambda_dist}). In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows. Additionally, while the Fellegi-Sunter model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence. 
\end{remark}

@@

Next, the full conditional for $\pi$ is given by
\begin{align}
	\pi^{(s+1)}|\bm{Z}^{(s)},  \alpha_{\pi}, \beta_{\pi}, \pi \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
\end{align}
where $n_{AB}(\bm{Z}^{(s)}) = \sum_{j=1}^{n_B} I(Z_j^{(s)} \leq n_A)$.

@@

The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_A  n_B$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors. One approach to reduce the number of comparisons is blocking, which places similar records into partitions, or ``blocks'' \citep{christen2019data}. In deterministic blocking, the modeler chooses fields thought to be highly reliable and only compares records that agree on those fields. The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains. Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.

@@

With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage. First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors. Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches.

@@

Since each component $\gamma_{ij}^f$ is discrete, there are only finitely many possible realizations of the comparison vector $\gamma_{ij}$. Let $P$ be the number of patterns realized in $\gamma$. It is bounded above by $P^{*} =  \prod_{f=1}^F (L_f + 1)$, where the addition of 1 to $L_f$ for each field accounts for the possibility of missing values. This quantity is determined by $F$ and $L_f$, and does not scale with $n_A$ or $n_B$. To obtain a memory efficient representation, we map the agreement pattern of each record pair to a unique integer. \cite{enamorado2019using} accomplished this through a hashing function, which we modify to explicitly handle missing values:
\begin{align}
	\label{eqn:hashing}
	h^{*}(\gamma_{ij}) = \sum_{f = 1}^F I_{obs}(\gamma_{ij}^f)2^{\gamma_{ij}^f + I(f>1)\sum_{d=1}^{f-1}(L_d)}.
\end{align}

We then map these integers to sequential integers $\{1, \ldots, P\}$. We refer to each unique agreement pattern as $h_p$, and to the set of unique agreement patterns as $\mathcal{P} = \{h_1, \ldots, h_P\}$. When the $(i,j)$ record pair exhibits agreement pattern $p$, we say $\gamma_{ij} = h_p$. In calculations, we will at times use the one-hot encoding of agreement pattern $h_p$, denoted $e(h_p)$. This is a vector of length $\sum_{f=1}^F L_f$ in which the $l + \sum_{k=1}^{f-1} L_k$ component is 1 when $\gamma_{ij}^f = l$, and 0 otherwise.

For example, with five fields with binary agreements and possible missingness, the number of possible patterns is bounded above by $P^{*} = 3^5 = 243$. Records $A_5$ and $B_7$ may exhibit agreement pattern $\gamma_{5,7} = (1, 1, 1, NA, 2)$, indicating exact agreement on the first three fields, missing information in the fourth field, and disagreement in the fifth field. Then (\ref{eqn:hashing}) gives $h^{*}(\gamma_{5,7}) = 2^1 + 2^3 + 2^5 + 0 + 2^{10} = 1066$. We map the unique values of $h^{*}(\gamma)$ to sequential integers. If 1066 is mapped to, for example, 42, then we would have $\gamma_{5,7} = h_{42}.$ This agreement pattern has the one hot encoding $e(h_{42}) = (1, 0, 1, 0, 1, 0, 0, 0, 0, 1)$.

@@

		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks for more manageable computations. Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.  Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B and B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.  For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records. For each $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.
